### 2021-05-24 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
moved SC paper into thesis as CH4 (replacing old version). Other edits.

The code that allows one to produce the numerical examples in this chapter is publicly available as a Julia package\footnote{https://github.com/snowball13/SphericalHarmonics.jl} that partners the ApproxFun package \cite{ApproxFun} ? however, this package is purely experimental at this stage.

@@

Before we proceed further, let's build up to our definition of the spherical harmonics. We first need to introduce a few classical orthogonal polynomials.

On the unit interval, $[-1,1]$, we note that there is a hierarchy of orthogonal polynomials (OPs) in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}:
\bseqn{
	& \ddx{x} P^{(a,b)}_l (x) = \half \: (l + a + b + 1) \: P^{(a+1,b+1)}_{l-1}(x) \\
	\implies & \dmdxm{x}{m} P_l(x) = {(l+m)! \over 2^m \: l!} \: P^{(m,m)}_{l-m}(x)
}
where $P^{(a,b)}_l (x)$ is the $l$ degree \textit{Jacobi polynomial}, and $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$. Jacobi polynomials are orthogonal with weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$
\bseqn{
	\int_{-1}^1 {P^{(a,b)}_l(x)}^2 \:(1-x)^a(1+x)^b \: \D x &=: \omega_{P,l}^{(a,b)} \\
	\int_{-1}^1 P^{(a,b)}_l(x) \: P^{(a,b)}_{l'}(x) \:(1-x)^a(1+x)^b \: \D x &= \omega_{P,l}^{(a,b)} \: \delta_{l, l'}.
}
Legendre polynomials are special cases of \textit{Legendre functions} when $l$ is a non-negative integer. Legendre functions are a class of univariate functions that are solutions to the differential equation
\bseqn{
	(1 - x^2) \dmdxm{x}{2}y - 2x \ddx{x} y + l(l+1)y = 0
}
which is helpfully known as the \textit{Legendre equation} \cite[14.2.1]{DLMF}. Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Legendre polynomial \cite[p5]{belousov1962tables}:
\bseqn{
	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = (-1)^m \; {(l+m)! \over 2^m \; l!} \; (1-x^2)^\frac{m}{2} \; P^{(m,m)}_{l-m}(x),
}
for $m = 0,1,2,\dots,l$. It is also standard to include a further definition for the case of $m$ negative -- that is we also define:
\bseqn{
	\Plm(x) &:= (-1)^m \; {(l+m)! \over (l-m)!} \; P_l^{\abs{m}}(x) = {(l+m)! \over 2^{\abs{m}} \; l!} \; (1-x^2)^{\abs{m} \over 2} \; \jaclmabs{l}{m}(x)
}
for $m = -l, \dots, -1$.

>>>>

The code that allows one to produce the numerical examples in this chapter is publicly available as a Julia package\footnote{https://github.com/snowball13/SphericalHarmonics.jl} that partners the ApproxFun package \cite{ApproxFun} -- however, this package is purely experimental at this stage.

@@

Before we proceed further, let's build up to our definition of the spherical harmonics. We first need to introduce a few classical orthogonal polynomials. \enquote{Orthogonal polynomials} (OPs) are defined as such by the condition that each polynomial is orthogonal to each other, with respect to a given inner product. In particular, they are orthogonal with respect to all lower degree polynomials.

On the unit interval, $[-1,1]$, we note that there is a hierarchy of OPs in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}:
\bseqn{
	& \ddx{x} P^{(a,b)}_l (x) = \half \: (l + a + b + 1) \: P^{(a+1,b+1)}_{l-1}(x) \\
	\implies & \dmdxm{x}{m} P_l(x) = {(l+m)! \over 2^m \: l!} \: P^{(m,m)}_{l-m}(x)
}
where $P^{(a,b)}_l (x)$ is the $l$ degree \textit{Jacobi polynomial}, and $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$. Jacobi polynomials are orthogonal with weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$
\bseqn{
	\int_{-1}^1 {P^{(a,b)}_l(x)}^2 \:(1-x)^a(1+x)^b \: \D x &=: \omega_{P,l}^{(a,b)} \\
	\int_{-1}^1 P^{(a,b)}_l(x) \: P^{(a,b)}_{l'}(x) \:(1-x)^a(1+x)^b \: \D x &= \omega_{P,l}^{(a,b)} \: \delta_{l, l'}.
}
Legendre polynomials are special cases of \textit{Legendre functions} when $l$ is a non-negative integer. Legendre functions are a class of univariate functions that are solutions to the differential equation
\bseqn{
	(1 - x^2) \dmdxm{x}{2}y - 2x \ddx{x} y + l(l+1)y = 0
}
which is helpfully known as the \textit{Legendre equation} \cite[14.2.1]{DLMF}. Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Legendre polynomial \cite[p5]{belousov1962tables}:
\bseqn{
	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = (-1)^m \; {(l+m)! \over 2^m \; l!} \; (1-x^2)^\frac{m}{2} \; P^{(m,m)}_{l-m}(x),
}
for $m = 0,1,2,\dots,l$. It is also standard to include a further definition for the case of $m$ negative -- that is we also define:
\bseqn{
	\Plm(x) &:= (-1)^m \; {(l+m)! \over (l-m)!} \; P_l^{\abs{m}}(x) = {(l+m)! \over 2^{\abs{m}} \; l!} \; (1-x^2)^{\abs{m} \over 2} \; \jaclmabs{l}{m}(x)
}
for $m = -l, \dots, -1$.

### 2021-05-24 papers/thesis/full/ch3-disk-slices/disk-slices.tex
moved SC paper into thesis as CH4 (replacing old version). Other edits.

\begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:construction}
	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \condref{trap} or \condref{disk} with $w_2$ being an even function hold.
$\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$. Then the 2D polynomials defined on $\Omega$
\bseqn{
	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
}
are orthogonal polynomials with respect to the weight \(W(x,y) := w_1(x) \: w_2\fpr(\frac{y}{\rho(x)}) \) on $\Omega$. 
\end{proposition}

@@

$\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight
\bseqn{
	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: \jacw^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
}
assuming that either \condref{trap} or \condref{disk} with  \(\jacw^{(a,b)}\) being an even function (i.e. $a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) =  (\delta-x^2)^a$) hold.
That is,
\bseqn{
	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
}
where for $f, g : \Omega \to \R$ the inner product is defined as 
\bseqn{
	\ip< f, \: g >_{W^{(a,b,c,d)}} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x.
}

@@

Finally, if \condref{trap} holds we have that 
$$
\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{m-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} = 0 \qfor m < n-1.
$$
If \condref{disk} holds we have that $\gamma^{(d,c)}_{k} = \gamma^{(c,c)}_{k} \equiv 0$ for any $k$.

>>>>

\begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:construction}
	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \bsrefcondition{cond:trap} or \bsrefcondition{cond:disk} with $w_2$ being an even function hold.
$\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$. Then the 2D polynomials defined on $\Omega$
\bseqn{
	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
}
are orthogonal polynomials with respect to the weight \(W(x,y) := w_1(x) \: w_2\fpr(\frac{y}{\rho(x)}) \) on $\Omega$. 
\end{proposition}

@@

$\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight
\bseqn{
	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: \jacw^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
}
assuming that either \bsrefcondition{cond:disk} or \bsrefcondition{cond:disk} with  \(\jacw^{(a,b)}\) being an even function (i.e. $a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) =  (\delta-x^2)^a$) hold.
That is,
\bseqn{
	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
}
where for $f, g : \Omega \to \R$ the inner product is defined as 
\bseqn{
	\ip< f, \: g >_{W^{(a,b,c,d)}} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x.
}

@@

Finally, if \bsrefcondition{cond:trap} holds we have that 
$$
\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{m-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} = 0 \qfor m < n-1.
$$
If \bsrefcondition{cond:disk} holds we have that $\gamma^{(d,c)}_{k} = \gamma^{(c,c)}_{k} \equiv 0$ for any $k$.
